Designing and Developing a Personalised Recommender System

### 2021-02-12 paper/paper.tex
finished

\author{\IEEEauthorblockN{lrfk99}
\IEEEauthorblockA{\textit{dept. Computer Science} \\
\textit{Durham University}\\
Durham, United Kingdom}
}

@@

\subsection{Related work review}
Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
are designed to combine the ability of KB techniques to give generic recommendations to new users with 
small profiles and the strength of CF to find peer users with unexpected shared preferences. 
This combination of techniques is particularly well suited to restaurants since a user may often be in the 
mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
type or restaurant, which is where the knowledge-based recommender is helpful. 
A collaborative-filtering algorithm will then be able to order all the restaurants which match this 
explicit criteria by how the user should rate them, based on the opinions of other similar users, 
helping them to discover new restaurants to try. 
These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
two techniques in the same cascaded style that I implement here.

@@

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
1 in the covid section. 
The $business.json$ file contains business data including location data, attributes, and cuisine. 
$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
that the review is written for. 
The file $user.json$ holds every user's first name along with other data and metadata which I don't use \cite{YelpDataset}. 
The data files $checkin.json$ and $tip.json$ are not used in my system. 
The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
preparation is required to save computation time and in order to provide more accurate predictions. 
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
cuisines in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for cuisines they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 cuisines because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
covid dataset removed. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).

@@

\subsection{Recommendation techniques/algorithms}
The first prediction method used in this hybrid system is a knowledge-based (KB) approach, which relies on knowledge 
about each of the restaurants based on its location and the type of food that it serves. 
This approach completely eliminates the cold-start problem, a new user will still receive relevant recommendations 
for the city they are in and the type of food they in the mood for when using a KB system. 
The second prediction method is a collaborative-filtering (CF) approach, specifically Single Value Decomposition (SVD) 
matrix factorisation. 
This approach was chosen because it scales well with large datasets, which is important since my dataset has 
roughly 120,000 entries. 
The CF technique here also introduces an element of variability in the recommendations, such that the user 
may not receive the same ordering of suggestions when they run the system multiple times with the same 
KB specifications, helping the user to try new things. 
This system allows the user to make ratings through the user interface, providing the ability to experiment with 
the system in terms of how new ratings affect the current user's recommendations as well as those of other similar users.

\subsection{Evaluation methods}
The evaluation of the system was done using offline experiments. 
The accuracy of ratings predictions was evaluated using Mean Absolute Error (MAE), which is calculated by the 
following equation:
\begin{equation}
    MAE = \frac{1}{|\mathcal{T}|} \sum_{(u,i)\in{\mathcal{T}}} {|\widehat{r}_{ui} - r_{ui}|}
\end{equation}
This metric is appropriate for this RS since it can handle large errors on a few items, which may occur in this system 
due to some rare skewed ratings. 
The MAE of this system is around 0.99.

@@

\subsection{Input interface}
The system has a command line based interface, so all user input is given via the command line. 
The system recognises the active user by their unique user ID. 
Some example IDs are provided in the 'README.txt' file for testing and demonstration purposes. 
Only explicit user data is gathered in order to make the system more explainable and transparent, 
and users are made aware of which data is collected, how it is collected, and for which purposes. 
This data includes the user's nearest city (of the 10 available) and the type of food that 
they are in the mood for. 
The user is able to change these choices from the main menu, as well as switch to a different user 
for testing purposes. 
When the user is presented with a list of recommendations, they can then choose one that they want 
to see more information about and from here they can leave a rating, return to the list, or return to 
the main menu. 
User profiles can be updated by giving a rating of a restaurant through the system. 
Any rating given will affect future recommendations given by the system to that user 
and to other similar users.

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
This minimisation is performed by stochastic gradient descent:
\begin{align}
    \begin{split}
    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
    \end{split}
\end{align}
where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}. 
The knowledge-based aspect of the algorithm produces recommendations based on the explicit location and cuisine 
preferences of the user. 
These are then cascaded down to the CF algorithm, which assigns each a predicted rating based on the ratings data. 
The 8 recommendations with the highest predicted rating are then presented to the user.

\subsection{Output interface}
Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
a restaurant to try, without there being too many options which might overwhelm the user. 
If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
would require the user to scroll around to read different parts of the output. 
Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
background will help accommodate users who are colourblind or who require a high-contrast interface. 
The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
requires the text to be larger then they can increase the text size of their terminal. 
Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
This, along with an option in the main menu to view information about the data the system uses and how it is used, 
further adds to the explainability and transparency. 
\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{output.png}
    \caption{System output}
    \label{fig:output}
\end{figure}

>>>>

\author{\IEEEauthorblockN{lrfk99}
\IEEEauthorblockA{\textit{dept. Computer Science} \\
\textit{Durham University} \\
Durham, United Kingdom}
}

@@

\subsection{Related work review}
Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
are designed to combine the ability of KB techniques to give appropriate recommendations to new users with 
the ability of CF to find peer users with unexpected shared preferences. 
This is appropriate for restaurants since a user may be in the mood for a specific cuisine and only want to see 
suggestions relevant to that, which is where the KB system comes in. 
The CF algorithm will then be able to sort these restaurants provided by the KB system by their predicted 
rating by the user, based on the opinions of other similar users, helping them to discover new restaurants to try. 
These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
two techniques in the same cascaded style that I implement here.

@@

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The $business.json$ file contains business data including location data, attributes, and cuisine. 
$review.json$ contains full review text data including the $user\_id$ of the review and the $business\_id$ 
that the review is for. 
The file $user.json$ holds every user's first name along with other data and metadata which I don't 
require \cite{YelpDataset}. 
The data files $checkin.json$, $tip.json$, and $photos.json$ are not used in my system. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
So I converted every file I used in the system into a typical JSON file, allowing for extremely fast 
data lookups in the system. 
I filtered the list of all Yelp businesses to restaurants and then further by selecting only those which were still 
open and fit into one or more of the top 12 cuisines in terms of number of restaurants because it's enough to provide 
a wide variety of options to the user without being so many that the user is overwhelmed. 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users, 
to ensure I was only using high quality reviews which others found helpful. 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as 
follows: ($User\_ID$, $Business\_ID$, $Rating$).

@@

\subsection{Recommendation techniques/algorithms}
The first prediction method used in this hybrid system is a knowledge-based (KB) approach, which relies on knowledge 
about each of the restaurants based on its location and the type of food that it serves. 
This approach completely eliminates the cold-start problem, a new user will still receive relevant recommendations 
for the city they are in and the type of food they in the mood for when using a KB system. 
The second prediction method is a collaborative-filtering (CF) approach, specifically Single Value Decomposition (SVD) 
matrix factorisation. 
This approach was chosen because it scales well with large datasets. 
The CF technique here also introduces an element of variability in the recommendations, such that the user 
may not receive the same ordering of suggestions when they run the system multiple times with the same 
KB specifications, helping the user to try new things. 
This system allows the user to make ratings through the user interface, providing the ability to experiment with 
the system in terms of how new ratings affect the current user's recommendations as well as those of other similar users.

\subsection{Evaluation methods}
The evaluation of the system was done using offline experiments. 
The accuracy of ratings predictions was evaluated using Mean Absolute Error (MAE), which is calculated by the 
following equation:
\begin{equation}
    MAE = \frac{1}{|\mathcal{T}|} \sum_{(u,i)\in{\mathcal{T}}} {|\widehat{r}_{ui} - r_{ui}|}
\end{equation}
This metric is appropriate for this RS since it can handle large errors on a few items, which may occur in this system 
due to some rare skewed ratings.

@@

\subsection{Input interface}
The system has a command line interface, and recognises the active user by their unique user ID. 
Some example IDs are provided for testing and demonstration purposes. 
Only explicit user data is gathered, making the system more explainable and transparent, 
and users are made aware of which data is collected, how it is collected, and for which purposes. 
This data includes the user's nearest city (of the 10 available) and the type of food that 
they are in the mood for. 
The user is able to change these choices from the main menu, as well as switch to a different user 
for testing purposes. 
When the user is presented with a list of recommendations, they can then choose one that they want 
to see more information about and leave a rating, which will update their future recommendations.

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error by stochastic gradient descent: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
The knowledge-based aspect of the algorithm produces recommendations based on the explicit location and cuisine 
preferences of the user. 
These are then cascaded down to the CF algorithm, which assigns each a predicted rating based on the ratings data. 
The 8 recommendations with the highest predicted rating are then presented to the user.

\subsection{Output interface}
Recommendations and prediction scores are presented to the user via the command line interface using an ASCII table. 
The top 8 recommendations are output since this is enough variety for the user to find something new to try, 
without being too overwhelming. 
Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
background will help accommodate users who are colourblind or who require a high-contrast interface. 
The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
requires the text to be larger then they can increase the text size of their terminal. 
Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
This, along with an option in the main menu to view information about the data the system collects and uses, 
further adds to the explainability and transparency. 
\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{output.png}
    \caption{System output}
    \label{fig:output}
\end{figure}
